Mining Historical Advertisements in Digitized Newspapers

### 2020-12-16 ms/article.tex
final revision

 \documentclass[USenglish]{article}
\usepackage{babel}[USenglish]
\usepackage[utf8]{inputenc}%(only for the pdftex engine)
%\RequirePackage[no-math]{fontspec}[2017/03/31]%(only for the luatex or the xetex engine)
\usepackage[small]{dgruyter}
\usepackage{wrapfig}
\usepackage{booktabs}
\captionsetup{justification=centering}
\usepackage{subcaption} 
\usepackage{microtype}
\usepackage{lmodern}% http://ctan.org/pkg/lm
%\usepackage[backend=bibtex]{biblatex}

@@

  \author*[1]{Melvin Wevers}
  %\author[2]{...}
  %\author[1]{Melvin Wevers} 
  \runningauthor{Wevers}
  \affil[1]{DHLAB - KNAW Humanities Cluster}
  %\affil[2]{...}
  \title{Mining Historical Advertisements in Digitized Newspapers}
  \runningtitle{Mining Historical Ads}
  %\subtitle{}
  \abstract{Historians have turned their focus to newspaper articles as a proxy of public discourse, while advertisements remain an understudied source of digitized information. This paper focuses on how historians can use computational methods to work with extensive collections of advertisements. First, this chapter analyzes metadata to understand better the different types of advertisements, which come in a wide range of shapes and sizes. Information on the size and position of advertisements can help to construct particular subsets of advertisements. Second, this chapter describes how textual information can be extracted from historical advertisements that can subsequently be used for historical analysis of trends and particularities. For this purpose, we present a case study based on cigarette advertisements.}
  \keywords{historical advertisements, text mining, digitized newspapers, digital history}
  %\classification[PACS]{...}
  %\communicated{...}
  %\dedication{...}
  \received{...}
  \accepted{...}
  \journalname{...}
  \journalyear{...}
  \journalvolume{..}
  \journalissue{..}
  \startpage{1}
  \aop
  \DOI{...}

@@

In recent years, we have seen an explosive growth of digitized historical newspapers. Innovations in natural language processing have extended the possibilities for historians to extract information from large corpora of digitized historical texts. National libraries and project such as impresso, Newseye, and Oceanic Exchanges offer access to digitized archives of historical newspapers.\footnote{\url{https://impresso-project.ch/}, \url{https://www.newseye.eu/}, \url{https://oceanicexchanges.org/}}

For many historians, newspapers provide a longitudinal understanding of public discourse~\cite{postman_amusing_2005,vree_nederlandse_1989}. Newspapers are not the only gateway to public discourse, since they do not capture public discourse in its entirety. In its function as a proxy, a newspaper operates as a transceiver; it is both the producer and the messenger of public discourse~\cite[17-18]{schudson_power_1982}. On a surface level, newspapers inform us about the views of journalists and people that were interviewed by these journalists. However, as Margaret Marshall claims, scholars can also uncover the ``values, assumptions, and concerns, and ways of thinking that were a part of the public discourse of that time'' by analyzing ``the arguments, language, the discourse practices that inhabit the pages of public magazines, newspapers, and early professional journals~\cite[8]{marshall_contesting_1995}.'' With easier access to and increased availability of newspaper repositories, studies of the representation of ideas, values, and practices in public discourse gained traction~\cite{van_eijnatten_eurocentric_2019,daems_workers_2019}.

Even though newspapers contain a considerable amount of advertisements, these remain an understudied source in computational studies of public discourse. This is surprising since advertisements are rich and varied carriers of information of the past. In his seminal work \textit{Advertising the American Dream}, Roland Marchand argues that adverts offer a lens on the past and provide ``insight into the ideals and aspirations of past realities [...] they show the state of technology, the social functions of products, and provide information on the society in which a product was sold~\cite{marchand_advertising_1985}.'' Others point out that even though adverts provide perspectives on the past, this is a distorted one, as the content of ads is driven by commercial interest~\cite{fox_mirror_1997}. 
%
Marchand acknowledges this criticism and conceptualizes advertisements as distorted mirrors. He argues that despite the primary function of advertisements to sell products, they still communicated social and cultural values, albeit in a somewhat distorted manner~\cite{marchand_advertising_1985,lears_fables_1994}.

Moreover, one could argue that for ads to be successful, they needed to resonate with their audience. As such, they had to be reflective of aspects of public discourse. Still, it is crucial to remain critical of the skewed representation of ideas and values in historical advertisements.

@@

The ability to extract information from large numbers of advertisements, both synchronically and diachronically, allows scholars to study cultural expressions on a macro-scale. However, the use of computational methods can also offer fine-grained contextualizations of specific expressions. More often than not, historical interpretations drawn from advertisements are based on a small selection of ads, which might lead to cherry-picking. The ability to chart trends over time makes it possible to model whether findings in smaller subsets can be generalized, or whether specific expressions deviate from general patterns. On an intermediate, mesoscale, one can also more easily find variations of a single cultural expression. These insights gained with computational methods supplement existing modes of analysis.

However, we also have to take into account that the quality of digitized historical sources is often sub-optimal, despite advances in Optical Character Recognition (OCR) software and natural language processing. It is especially challenging for OCR software to recognize text in advertisements correctly. In addition to more generic factors such as the quality of paper or the printing technique, text in advertisements is often presented in varying sizes and fonts or as part of a heavily-stylized logo. OCR software regularly turns these forms of textual content into gibberish. The suboptimal text quality makes it more challenging to study adverts than articles.

Despite these shortcomings, we can still extract meaningful information from the OCR-ed text. Given the amount of data, we can still find patterns in advertising discourse despite the noisy nature of the text. In this process, we have to be selective of the methods that we use, since their performance can be impacted in different ways by imperfect OCR~\cite{van_strien_assessing_2020,hillQuantifyingImpactDirty2019}. In addition to the OCR-ed text, we can also learn about trends in advertising from metadata on the position and size of advertisements in newspapers. We can, thus, study multiple aspects of advertisements using computational means.

This chapter showcases how we can use computational techniques to study advertisements at scale in extensive collections of digitized newspapers. The first section shows how we can use metadata to examine trends in advertising. The second section gives examples of how text mining can be used to extract information from advertisements. This step is explained by demonstrating a case on product nationalities associated with cigarettes. In this case study, we rely on text mining to better understand changes and continuities in the associations with nationalities in advertisements for cigarettes.

\section{Metadata Analysis: Advertisements in Almost All Shapes and Sizes}

The digitization of newspapers has made it possible to study advertisements at scale using key word searches. However, not all digitized newspaper collections have segmented the articles and advertisements, presenting users with full-page scans that contain multiple document types, for example, advertisements and articles. In cases where text blocks were segmented but not classified, text classifiers can be used to identify document types~\cite{bilgin_utilizing_2018}. After training a text classifier with annotated data, the classifier could then predict metadata for unseen new data. Recently, there have been efforts to rely on computer vision to segment and classify the elements on a newspaper page (e.g. articles, tables, ads).\footnote{See, for example, Ben Lee's project \textit{Newspaper Navigator} at the Library of Congress. \url{https://github.com/LibraryOfCongress/newspaper-navigator}}

Fortunately, Delpher, the digitized Dutch newspaper archive hosted by the National Library of the Netherlands (KB), includes segmented documents and metadata on the document type.\footnote{\url{http://www.delpher.nl}. The metadata on document type was manually added.} Access to this type of metadata  allows researchers to filter for advertisements. For the twentieth century alone, we can quickly assess that Delpher holds over thirty million advertisements in national and regional newspapers. Browsing through the results, it becomes clear that advertisements come in a wide range of shapes and sizes. We can find tiny, square-shaped advertisements;  column-shaped classifieds advertisements; and full-age spreads (see Figure~\ref{fig:classifieds} for an example of a column of classified ads). The content and target audiences of these ads vary considerably. How can we deal with this variation, without more specific metadata information? 
\begin{wrapfigure}{r}{0.15\textwidth}
  \vspace{-10mm}
  \begin{center}
    \includegraphics[height=0.40\textheight]{figures/classifieds}
  \end{center}
  \caption{Classified adverts from \textit{De Tijd}, 19 January, 1957}
  \vspace{-7mm}
  \label{fig:classifieds}
\end{wrapfigure}
%
This section shows how we can use metadata on the size and position of advertisements to cluster types of advertisements and subsequently filter out subsets. The analysis in this section is based on the metadata of the national newspaper \textit{Trouw} (1946-1995).\footnote{For most of the examples in this chapter, direct access to the metadata and newspaper data is required. Researchers can contact KB for API access to the newspapers. Code is available on: \url{https://github.com/melvinwevers/eldorado}} An initial exploratory data analysis shows that the title contains instances of bad segmentation, in which small parts of advertisements appeared as separate advertisements. To exclude these segmentation errors, we filtered out ads with a width or height smaller than 100 pixels. After removing these advertisements, \textit{Trouw} contains about 1.31 million advertisements for the period 1945-1995. From Figure~\ref{fig:ad_curve}, we can gauge that the total number of advertisements increased between 1946 and 1995, with a sudden increase in the early 1980s. Alongside this increase in the number of ads, we also see a sudden, albeit slight decrease in the size (Figure~\ref{fig:mean_size}). We plotted took the $log_{10}$ of the mean size to visualize the rate of change better.

\begin{figure}%
  \centering
  %
  %\qquad
  \includegraphics[width=.9\textwidth]{figures/Trouw_ad_curve}%
  \caption{Yearly number of advertisements in \textit{Trouw}. Green triangles indicate peaks and red triangle troughs.}%
  \label{fig:ad_curve}%
\end{figure}

\begin{figure}%
  \centering
  %
  %\qquad
  \includegraphics[width=.9\textwidth]{figures/mean_size}%
  \caption{Mean size of advertisements in \textit{Trouw}}
  \label{fig:mean_size}%
\end{figure}

In addition to their size, we can also examine where ads are positioned in the newspaper and on the individual pages. Using the width and height information of advertisements, we can create a heat map that shows which parts of the page are taken up by advertisements. There is a clear difference in pixel density between odd and even pages in \textit{Trouw} (Figure~\ref{fig:heatmapa} \& \ref{fig:heatmapb}). This can be explained by the fact that advertising on odd pages is more expensive than on even pages, as well as the common practice that new sections generally start on the odd pages. Moreover, we see that on the even pages, the upper-left corner, which is closer to the margin, is less populated than the lower-right side.

To get a better grasp of the different groups of ads based on their 
size and their position in the newspaper, we can also apply unsupervised clustering methods, e.g. Gaussian Mixture modeling. We base this clustering on the features: width, height, and relative position of the ad in the issue. The latter indicates how close the ad is to the front page or the last page. The distribution of these features contains multiple peaks, indicative of multiple sub-distributions. To be able to capture the distributions of these subpopulations, we can use Gaussian Mixture modeling, which can estimate the parameters of these mixtures.\footnote{Scikit-Learn offers the Gaussian mixture algorithm for the estimation of a mixture model.} Compared to K-means clustering, mixture models incorporate information on the covariance structure of the data, making it possible to capture clusters with varying distributional shapes. We apply the clustering separately to even and odd pages, since there seem to be different generative principles at play, as evinced by the heat maps in Figure~\ref{fig:heatmapa} \& \ref{fig:heatmapb}. After estimating the optimal number of clusters for even and odd pages ($n$ = 8 and $n$ = 7), we fit a Gaussian Mixture model to the data.

@@

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.8\textwidth}
    \includegraphics[width=\textwidth]{figures/even_sizes}
    \caption{Clusters of advertisement based on width, height and relative page number on even pages in \textit{Trouw}}
    \label{fig:clustera}
  \end{subfigure}
  %
  \begin{subfigure}[b]{0.8\textwidth}
    \includegraphics[width=\textwidth]{figures/odd_sizes}
    \caption{Clusters of advertisement based on width, height and relative page number on odd pages in \textit{Trouw}}
    \label{fig:clusterb}
  \end{subfigure}
  \caption{Cluster analysis of advertisements in \textit{Trouw}.
\end{figure}

The results of this clustering are presented in Figures~\ref{fig:clustera} \& ~\ref{fig:clusterb}. These plots are based on a sample of 50,000 ads from even and odd pages. We see apparent differences in size and position in the newspaper between the odd and even pages. The odd pages are less structures and lack a clear signal and show more noise, especially for ads with larger sizes. The ads on the even pages seems to be more structured. For some of these clusters, there is more variation, while others are closely clustered. A noteworthy cluster on the even pages is the brown cluster, which refers to full-page ads. 
%
We can also see that these ads appear toward the end of the newspaper. Newspapers are generally structured in columns, which is visible in the distinct structure present in the widths of advertisements (x-axis). The height of ads (y-axis) has more variance than the width, nevertheless, the algorithm distinguishes clusters based on their height. In terms of relative position, the number of advertisements increases towards the back of the newspaper (z-axis). Ads with a small width and a more considerable height, possibly classifieds ads (see Figure~\ref{fig:classifieds}, also appear toward the end of the paper.

The information contained in these clusters can help us filter out particular subsets of advertisements. For example, maybe we are not interested in classifieds ads or tiny ads. This method can help us determine the boundaries of these ads in terms of size and position.

We can also engineer additional features that help us refine our subsets even further. A useful feature, for instance, is character proportion, which can be calculated by dividing the number of characters by the size (width $x$ height) of the adverts. We can use this metric to filter out textual advertisements or those that predominantly consist of visual material. A similar feature is the ratio of digits to characters. This feature can be used to distinguish ads that mostly list prices of imported goods or wholesale goods, from other ads. 
%
Ostensibly, these numbers are dependent on the quality of the text extracted by Optical Character Recognition (OCR). Notwithstanding variations between newspapers and differences over time, these numbers allow us to filter particular types of advertisements. For example, we can filter out ads by inspecting particular clusters in the distribution of the character proportion. This form of exploratory data analysis can, for example, help to quickly filter out specific types of advertisements, such as classifieds. These advertisements are characterized by a higher than average height and smaller width, as well as a higher character proportion. We could also combine this metadata information with keyword search to further filter out types of ads for specific products.

In this section, we have demonstrated a possible approach to using metadata to learn more about the structure of advertisements in historical newspapers. There are many more directions to explore. However, this section underscores that metadata should not be overlooked and can be used in conjunction with text analysis to guide further analysis.

@@

From this corpus, we extracted a sub-corpus of advertisements that contain the three most common spelling variants of cigarette: \textit{cigaret}, \textit{sigaret}, and \textit{cigarette}. These singular and plural variants are queried using the following regular expression: 
\verb/'cigaret*\w+', 'sigaret*\w+'./ This query yielded 43,781 advertisements. Figure~\ref{fig:cigarette_ad_curve} displays the distribution of the relative number of these cigarette advertisements. The trend line shows that the relative number of cigarette advertisements grew until the early 1920s, after which it decreased. In the 1950s, the relative number of ads again peaked, and after the 1960s, the number of advertisements that included variants of the word cigarette dropped considerably, suggesting that after the 1960s, cigarette manufacturers advertised less in Dutch newspapers.

\begin{figure}%
  \centering
  \includegraphics[width=.9\textwidth]{figures/cigarette_ads}%
  \caption{Relative Number of Cigarette Advertisements 1890-1990}%
  \label{fig:cigarette_ad_curve}%
\end{figure}

Plotting time series gives an overview of the temporal distribution of advertisements for a particular product. However, the search terms should be carefully selected. Using the different spelling variations, captured with the regular expression, this query covers most of the historical variations in the twentieth century. In the case of cigarettes, the decrease in the last quarter of the twentieth century can in part be explained by the fact that advertisers stopped referencing cigarettes explicitly. Rather, they turned to ads that include brand names or logos but not 'cigarette'. In such instances, computer vision could help to detect cigarettes or the occurrence of particular logos in advertisements. Many of these brand names are ambiguous, and querying them could also return advertisements for unrelated products. 
%
Moreover, the period during the Second World War should also be scrutinized more closely, since only a few newspapers were published throughout this period, and the size and content of newspapers transformed drastically. Even when using relative frequencies to offset the changing number of newspapers, we cannot account for such factors.

\subsection{Cigarettes and Product Nationalities}
While text mining allows to study of large-scale trends, it can also help to explore more specific ways in which advertisers framed a product. One type of framing is the nationality of a consumer good, also known as product nationality. The relationship between products and constructions of nationality has long since been debated in consumer and marketing studies~\cite{thakor_brand_1996, hull_cultural_2016}. A key finding in these studies is that a strong connection to a favored country persuades consumers to spend more money on products with such a connection~\cite{menapace_consumers_2011}. The representations of product nationality are more than mere associations, they are ``powerful narratives about the meanings and values transferred by products from their origin to their destination~\cite{askegaard_product-country_1998}.''

@@

\begin{figure}
  \centering
  \includegraphics[width=.9\textwidth]{figures/example_buffalo.jpg}
  \caption{Advertisements for Buffalo cigarettes in \textit{Limburger Koerier}, April 13, 1938}
  \label{fig:buffalo_example}
\end{figure}

The advertisement in Figure~\ref{fig:buffalo_example} is a clear example of the association between product nationality and particular characteristics. This  1938 advert for the Buffalo cigarette linked the brand to the United States in several ways. First, the brand's name Buffalo denoted the emblematic American prairie animal as well as the city of Buffalo in upstate New York. The relationship between the brand and the United States was further enforced by a small print mentioning its producer: The Cumberland Company from Clarksville, USA.
%
Second, in addition to these textual cues, the advertisement included a visual signifier: a background image of a giant cowboy bending over a Dutch tulip field. This picture of a cowboy—an exemplar of American culture—further substantiated Buffalo as an American cigarette. Furthermore, this image expressed the towering dominance of American products in the Netherlands.
%
Third, the ad presented the Buffalo cigarette as having an American product nationality by describing it as ``the tastiest and spiciest American cigarette.'' The geographical association to the United States suggested the product’s country of origin, but also signified a particular taste specific to American techniques of tobacco preparation.

The Buffalo ad is of course just one example. How does this particular advertisement compare to others, and is its messaging specific to cigarettes with an American product nationality? Using text mining, we can gather the information that help us answer questions like these.

>>>>

\documentclass[USenglish]{article}
\usepackage{babel}[USenglish]
%\usepackage[utf8]{inputenc}%(only for the pdftex engine)
\RequirePackage[no-math]{fontspec}[2017/03/31]%(only for the luatex or the xetex engine)
\usepackage[small]{dgruyter}
\usepackage{wrapfig}
\usepackage{booktabs}
\captionsetup{justification=centering}
\usepackage{subcaption} 
\usepackage{microtype}
\usepackage{lmodern}% http://ctan.org/pkg/lm
%\usepackage[backend=bibtex]{biblatex}

@@

  \author*[1]{Melvin Wevers}
  \runningauthor{Wevers}
  \affil[1]{Department of History, University of Amsterdam, Amsterdam, the Netherlands}

  \title{Mining Historical Advertisements in Digitized Newspapers}
  \runningtitle{Mining Historical Ads}
  %\subtitle{}
  \abstract{Historians have turned their focus to newspaper articles as a proxy of public discourse, while advertisements remain an understudied source of digitized information. This paper shows how historians can use computational methods to work with extensive collections of advertisements. First, this chapter analyzes metadata to better understand the different types of advertisements, which come in a wide range of shapes and sizes. Information on the size and position of advertisements can be used to construct particular subsets of advertisements. Second, this chapter describes how textual information can be extracted from historical advertisements, which can subsequently be used for a historical analysis of trends and particularities. For this purpose, we present a case study based on cigarette advertisements.}
  \keywords{historical advertisements, text mining, digitized newspapers, digital history}
  %\classification[PACS]{...}
  %\communicated{...}
  %\dedication{...}
  \received{...}
  \accepted{...}
  \journalname{...}
  \journalyear{...}
  \journalvolume{..}
  \journalissue{..}
  \startpage{1}
  \aop
  \DOI{...}

@@

In recent years, we have seen an explosive growth of digitized historical newspapers. Innovations in natural language processing have extended the possibilities for historians to extract information from large corpora of digitized historical texts. National libraries and projects such as Impresso, Newseye, and Oceanic Exchanges offer access to digitized archives of historical newspapers.\footnote{\url{https://impresso-project.ch/}, \url{https://www.newseye.eu/}, \url{https://oceanicexchanges.org/}}

For historians, newspapers provide a longitudinal understanding of public discourse~\cite{postman_amusing_2005,vree_nederlandse_1989}. Newspapers are not the only gateway to public discourse, since they do not capture public discourse in its entirety. In its function as a proxy, a newspaper operates as a transceiver; it is both the producer and the messenger of public discourse~\cite[17-18]{schudson_power_1982}. On a surface level, newspapers inform us about the views of journalists and people that were interviewed by these journalists. However, as Margaret Marshall claims, scholars can also uncover the ``values, assumptions, and concerns, and ways of thinking that were a part of the public discourse of that time'' by analyzing ``the arguments, language, the discourse practices that inhabit the pages of public magazines, newspapers, and early professional journals~\cite[8]{marshall_contesting_1995}.'' With easier access to and increased availability of newspaper repositories, studies of the representation of ideas, values, and practices in public discourse have gained traction~\cite{van_eijnatten_eurocentric_2019,daems_workers_2019}.

Even though newspapers contain a considerable amount of advertisements, these remain an understudied source in computational studies of public discourse. This is surprising since advertisements are rich and varied carriers of information of the past. In his seminal work \textit{Advertising the American Dream}, Roland Marchand argues that adverts offer a lens on the past and provide ``insight into the ideals and aspirations of past realities [...] they show the state of technology, the social functions of products, and provide information on the society in which a product was sold~\cite{marchand_advertising_1985}.'' Others point out that even though adverts provide perspectives on the past, this is a distorted one, as the content of ads is driven by commercial interest~\cite{fox_mirror_1997}. 
%
Marchand acknowledges this criticism and conceptualizes advertisements as distorted mirrors. He argues that despite the primary function of advertisements to sell products, they still communicated social and cultural values, albeit in a somewhat distorted manner~\cite{marchand_advertising_1985,lears_fables_1994}. Moreover, one could argue that for ads to be successful, they needed to resonate with their audience. As such, they had to be reflective of aspects of public discourse. Still, it is crucial to remain critical of the skewed representation of ideas and values in historical advertisements.

@@

The ability to extract information from large numbers of advertisements, both synchronically and diachronically, allows scholars to study cultural expressions on a macro-scale. This process has also been described as distant reading~\cite{moretti_distant_2013}. In this chapter, I refrain from using this polemic term, which is often used in opposition to close reading. In this binary opposition, quantitative approaches are incorrectly equated with distant reading and qualitative ones as a form of close reading. When switching between perspectives of analysis, there is more than merely the binary of close and distant. In my view, the use of computational methods can also offer fine-grained contextualized interpretations of particular expressions. More often than not, historical interpretations drawn from advertisements are based on a small selection of ads, which opens the door to cherry-picking. The ability to chart trends over time makes it possible to model whether findings in smaller subsets can be generalized, or whether specific expressions deviate from general patterns. On an intermediate, mesoscale, one can also more easily find variations of a single cultural expression.

While computational methods open up new modes of analysis of historical sources, we have to take into account that the quality of these digitized materials is often sub-optimal, despite continuous advances in Optical Character Recognition (OCR) software and natural language processing. It is especially challenging for OCR software to recognize text in advertisements correctly. In addition to more generic factors such as the quality of paper or the printing technique, text in advertisements is often presented in varying sizes and fonts or as part of a heavily-stylized logo. OCR software regularly turns these forms of textual content into gibberish. The suboptimal text quality makes it more challenging to study adverts than articles.

Despite these shortcomings, we can still extract meaningful information and patterns from the OCR-ed text. In this process, we have to be selective of the methods that we use, since their performance can be impacted in different ways by imperfect OCR~\cite{van_strien_assessing_2020,hillQuantifyingImpactDirty2019}. In addition to the OCR-ed text, we can also learn about trends in advertising from metadata on the position and size of advertisements in newspapers. We can, thus, study multiple aspects of advertisements using computational means.

This chapter showcases how we can use computational techniques to study advertisements at scale in digitized newspapers. The first section shows how we can use metadata to examine trends in advertising. The second section gives examples of how text mining can be used to extract information from advertisements. This step is explained by demonstrating a case on product nationalities associated with cigarettes. In this case study, we rely on text mining to better understand changes and continuities in the associations with nationalities in advertisements for cigarettes.

\section{Metadata Analysis: Advertisements in almost all Shapes and Sizes}

\begin{wrapfigure}{r}{0.25\textwidth}
  \begin{center}
    \includegraphics[height=0.40\textheight]{figures/classifieds}
  \end{center}
  \caption{Classified adverts from \textit{De Tijd}, 19 January, 1957}
  \label{fig:classifieds}
\end{wrapfigure}

The digitization of newspapers has made it possible to study advertisements at scale using keyword searches. However, not all digitized newspaper collections have segmented the articles and advertisements, presenting users with full-page scans that contain multiple document types, for example, advertisements and articles. In cases where text blocks were segmented but not classified, text classifiers can be used to identify document types~\cite{bilgin_utilizing_2018}. After training a text classifier with annotated data, the classifier can then predict metadata for unseen new data. Recently, there have been efforts to rely on computer vision to segment and classify the elements on a newspaper page (e.g. articles, tables, ads).\footnote{See, for example, Ben Lee's project \textit{Newspaper Navigator} at the Library of Congress. \url{https://github.com/LibraryOfCongress/newspaper-navigator}}

Fortunately, Delpher, the digitized Dutch newspaper archive hosted by the National Library of the Netherlands (KB), includes segmented documents and metadata on the document type.\footnote{\url{http://www.delpher.nl}. The metadata on document type was added by manual annotators.} Access to this type of metadata allows researchers to filter for advertisements. For the twentieth century alone, we can quickly assess that Delpher holds over thirty million advertisements in national and regional newspapers. Browsing through the results, it quickly becomes clear that advertisements come in a wide range of shapes and sizes. We can find tiny, square-shaped advertisements;  column-shaped classifieds advertisements; and full-age spreads (see Figure~\ref{fig:classifieds} for an example of a column of classified ads). The content and target audiences of these ads vary considerably. How can we deal with this variation, without more specific metadata information?

%
This section explains how we can use metadata on the size and position of advertisements to cluster types of advertisements and subsequently filter out subsets. The analysis in this section is based on the metadata of the national newspaper \textit{Trouw} (1946-1995).\footnote{For most of the examples in this chapter, direct access to the metadata and newspaper data is required. Researchers can contact KB for API access to the newspapers. Code is available on: \url{https://github.com/melvinwevers/eldorado}} An initial exploratory data analysis shows that the title contains instances of bad segmentation, in which small parts of advertisements appeared as separate advertisements. To exclude these segmentation errors, we filtered out ads with a width or height smaller than 100 pixels. After removing these advertisements, \textit{Trouw} contains about 1.31 million advertisements for the period 1945-1995. From Figure~\ref{fig:ad_curve}, we can gauge that the total number of advertisements increased between 1946 and 1995, with a sudden increase in the early 1980s. Alongside this increase in the number of ads, we also see a sudden, albeit slight decrease in the size (Figure~\ref{fig:mean_size}).\footnote{We log-transformed the mean size to visualize the rate of change better.}

\begin{figure}%
  \centering
  %
  %\qquad
  \includegraphics[width=\textwidth]{figures/Trouw_ad_curve}%
  \caption{Yearly number of advertisements in \textit{Trouw}. Green triangles indicate peaks and red triangle troughs.}%
  \label{fig:ad_curve}%
\end{figure}

\begin{figure}%
  \centering
  %
  %\qquad
  \includegraphics[width=\textwidth]{figures/mean_size}%
  \caption{Mean size of advertisements in \textit{Trouw}}
  \label{fig:mean_size}%
\end{figure}

In addition to their size, we can also examine where ads are positioned in the newspaper and where they can be found on the individual pages. Using the width and height information of advertisements, we created a heat map that shows which parts of the page are taken up by advertisements. There is a clear difference in pixel density between odd and even pages in \textit{Trouw} (Figure~\ref{fig:heatmapa} \& \ref{fig:heatmapb}). This can be explained by the fact that advertising on odd pages is more expensive than on even pages, as well as the common practice that new sections generally start on the odd pages. Moreover, we see that on the even pages, the upper-left corner, which is closer to the margin, is less populated than the lower-right side.

To get a better grasp of the groups of ads based on their size and their position in the newspaper, we can turn to unsupervised clustering methods. These methods can detect groups or clusters within a data set based on a set of features. We base this clustering on the features: width, height, and relative position of the ad in the issue. The latter indicates how close the ad is to the front page or the last page. The distribution of these features contains multiple peaks, indicative of multiple sub-distributions. To be able to capture the distributions of these subpopulations, we use Gaussian Mixture modeling, which can estimate the parameters of these mixtures.\footnote{Scikit-Learn offers the Gaussian mixture algorithm for the estimation of a mixture model.} Compared to K-means clustering, mixture models incorporate information on the covariance structure of the data, making it possible to capture clusters with varying distributional shapes. We apply the clustering separately to even and odd pages, since there seem to be different generative principles at play, as evinced by the heat maps in Figure~\ref{fig:heatmapa} \& \ref{fig:heatmapb}. After estimating the optimal number of clusters for even and odd pages ($n$ = 8 and $n$ = 7), we fit a Gaussian Mixture model to the data.

@@

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.8\textwidth}
    \includegraphics[width=\textwidth]{figures/even_sizes}
    \caption{Clusters of advertisement based on width, height and relative page number on even pages in \textit{Trouw}}
    \label{fig:clustera}
  \end{subfigure}
  %
  \begin{subfigure}[b]{0.8\textwidth}
    \includegraphics[width=\textwidth]{figures/odd_sizes}
    \caption{Clusters of advertisement based on width, height and relative page number on odd pages in \textit{Trouw}}
    \label{fig:clusterb}
  \end{subfigure}
  \caption{Cluster analysis of advertisements in \textit{Trouw}}.
\end{figure}

The results of this clustering are presented in Figures~\ref{fig:clustera} \& ~\ref{fig:clusterb}. These plots are based on a sample of 50,000 ads from even and odd pages. We see apparent differences in size and position in the newspaper between the odd and even pages. The odd pages are less structured and lack a clear signal and show more noise, especially for ads with larger sizes. The ads on the even pages seem to be more structured. For some of these clusters, there is more variation, while others are closely clustered. A noteworthy cluster on the even pages is the brown cluster, which refers to full-page ads. 
%
We can also see that these ads appear toward the end of the newspaper. Newspapers are generally structured in columns, which is visible in the distinct structure present in the widths of advertisements (x-axis). The height of ads (y-axis) has more variance than the width, nevertheless, the algorithm distinguishes clusters based on their height. In terms of relative position, the number of advertisements increases towards the back of the newspaper (z-axis). Ads with a small width and a more considerable height, possibly classifieds ads (see Figure~\ref{fig:classifieds} for an example), also appear toward the end of the paper.

The information contained in these clusters can help us focus on or filter out particular subsets of advertisements. This method can help us determine the boundaries of these ads in terms of size and position.

We can also engineer additional features that help us refine our subsets even further. A useful feature, for instance, is character proportion, which can be calculated by dividing the number of characters by the advert's size (width $x$ height) of the adverts. We can use this metric to filter out textual advertisements or those that predominantly consist of visual material. A similar feature is the ratio of digits to characters. This feature can be used to distinguish ads that mostly list prices of imported goods or wholesale goods, from other ads. 
%
Ostensibly, these numbers are dependent on the quality of the text extracted by Optical Character Recognition (OCR). Notwithstanding variations between newspapers and differences over time, these metrics allow us to filter particular types of advertisements. For example, we can filter out ads by inspecting particular clusters in the distribution of the character proportion. This form of exploratory data analysis can, for example, help to quickly filter out specific types of advertisements, such as classifieds. These advertisements are characterized by a higher than average height and smaller width, as well as a higher character proportion. We could also combine this metadata information with keyword searches to further filter out particular types of ads for selected products.

In this section, we have demonstrated a possible approach to using metadata to learn more about the structure of advertisements in historical newspapers. There are many more directions to explore. However, this section underscores that metadata should not be overlooked and can be used in conjunction with text analysis to guide and support further analysis.

@@

From this corpus, we extracted a sub-corpus of advertisements that contain the three most common spelling variants of cigarette: `cigaret', `sigaret', and `cigarette'. These singular and plural variants are queried using the following regular expression: 
\verb/'cigaret*\w+', 'sigaret*\w+'./ This query yielded 43,781 advertisements. Figure~\ref{fig:cigarette_ad_curve} displays the distribution of the relative number of these cigarette advertisements. The trend line shows that the relative number of cigarette advertisements grew until the early 1920s, after which it decreased. In the 1950s, the relative number of ads again peaked, and after the 1960s, the number of advertisements that included variants of `cigarette' dropped considerably, suggesting that after the 1960s, cigarette manufacturers advertised less in Dutch newspapers.

\begin{figure}%
  \centering
  \includegraphics[width=\textwidth]{figures/cigarette_ads}%
  \caption{Relative Number of Cigarette Advertisements 1890-1990}%
  \label{fig:cigarette_ad_curve}%
\end{figure}

Plotting time series gives an overview of the temporal distribution of advertisements for a particular product. However, the search terms should be carefully selected. Using the different spelling variations, captured with the regular expression, this query covers most of the historical variations in the twentieth century. In the case of cigarettes, the decrease in the last quarter of the twentieth century can in part be explained by the fact that advertisers stopped referencing cigarettes explicitly. Rather, they turned to ads that include brand names or logos but not 'cigarette'. In such instances, computer vision could help to detect depictions of cigarettes or the occurrence of particular logos in advertisements. Many of these brand names are ambiguous, and querying them could also return advertisements for unrelated products. 
%
Moreover, the period during the Second World War should also be scrutinized more closely, since the newspaper landscape differs considerably and the size and content of newspapers transformed drastically. Even when using relative frequencies to offset the changing number of newspapers, we cannot account for such factors.

\subsection{Cigarettes and Product Nationalities}
While text mining allows studying large-scale trends, it can also help to explore more specific ways in which advertisers framed a product. One type of framing is the nationality of a consumer good, also known as product nationality. The relationship between products and constructions of nationality has long since been debated in consumer and marketing studies~\cite{thakor_brand_1996, hull_cultural_2016}. A key finding in these studies is that a strong connection to a favored country persuades consumers to spend more money on products with such a connection~\cite{menapace_consumers_2011}. The representations of product nationality are more than mere associations, they are ``powerful narratives about the meanings and values transferred by products from their origin to their destination~\cite{askegaard_product-country_1998}.''

@@

\begin{figure}
  \centering
  \includegraphics[width=.5\textwidth]{figures/example_buffalo.jpg}
  \caption{Advertisements for Buffalo cigarettes in \textit{Limburger Koerier}, April 13, 1938}
  \label{fig:buffalo_example}
\end{figure}

The advertisement in Figure~\ref{fig:buffalo_example} is a clear example of the association between product nationality and particular characteristics. This  1938 advert for the Buffalo cigarette linked the brand to the United States in several ways. First, the brand's name Buffalo denoted the emblematic American prairie animal as well as the city of Buffalo in upstate New York. The relationship between the brand and the United States was further enforced by a small print mentioning its producer: The Cumberland Company from Clarksville, USA.
%
Second, in addition to these textual cues, the advertisement included a visual signifier: a background image of a giant cowboy bending over a Dutch tulip field. This picture of a cowboy---an exemplar of American culture---further substantiated Buffalo as an American cigarette. Furthermore, this image expressed the towering dominance of American products in the Netherlands.
%
Third, the ad presented the Buffalo cigarette as having an American product nationality by describing it as ``the tastiest and spiciest American cigarette.'' The geographical association to the United States suggested the product’s country of origin, but also signified a particular taste specific to American techniques of tobacco preparation.

The Buffalo ad is of course just one example. How does this particular advertisement compare to others, and is its messaging specific to cigarettes with an American product nationality? Using text mining, we can gather the information that helps us answer questions like these.
